---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# facts

<!-- badges: start -->
<!-- badges: end -->

The goal of facts is to simplify how the `factor` class is created, modified, and coerced.
`[base::factor()]` contains added sugar for recoding values and managing multiple constraints.
`facts()` takes a simplified approach and moves these nice-to-haves to separate functions.

## Todos

- [ ] Use default `NA` label (`"(null)"`) or something
- [ ] Be consistent about `values()`
- [ ] Bench mark a bit more
- [ ] Determine best way to adjust `values` (and `labels`) (e.g., `values<-()`) and special cases of `range`
    - `values(f, "levels") <- new_levels` for `character`
    - `values(f, "range") <- values_range` for `integer` and `Date`
    - `values(f, "levels", "remove") <- bad_values` for `character`
    - `drop_levels(f)` for undoing these things
    - `range_levels(f)` should add add a range for `integer` and `Date`

## Installation

You can install the development version of facts like so:

``` r
remotes::install_github("jmbarbone/facts")
```

## Creating facts

This is a basic example which shows you how to solve a common problem:

```{r example}
library(facts)
```

A `facts` object looks very much like a `factor` object, with a few differences.

```{r}
f1 <- factor(c(1, NA))
f2 <- facts(c(1, NA))

f1
f2

str(f1)
str(f2)

f2
levels(f2) <- c(1, "(null)")
f2
```


In general, `factor()` will always sort `x` and by default doesn't add `NA` values to levels.
`facts()`, however, will not sort character vectors, will use a reverse sort of logical, sorts numeric accordingly, and always includes `NA` labels when present in `x`.

```{r}
x <- c("a", "c", NA, "a", "b", NA, "a", "c")
factor(x) # sorted, NA removed by default
facts(x)   # unsorted, NA retained by default

x <- c(-1, 5, 2, NA, 3)
factor(x) # sorted
facts(x)   # sorted, but NA retained by default

x <- c(NA, FALSE, TRUE, FALSE, TRUE, NA)
factor(x)
facts(x)   # Sorted TRUE, FALSE, NA
```

## Modifying facts

One benefit of `factor()` is that you an recode on the fly.
This adds additional overhead to the function because there is a necessity in checking the `levels` and `labels` and matching up appropriate values.
The philosophy taken in `{facts}` is that the additional work of modifying the object should take place prior to `facts()`.
However, some cases, these sort of manipulations can benefit from reducing the number of unique values.

```{r}
x <- c("blue", "green", "red", "purple", "black", "white")
x <- sample(x, 100, TRUE)
id <- pseudo_id(x)
id
```

